VM-Centric Snapshot Deduplication for Converged Cloud Architectures

### 2014-07-19 snapshot_system/paper_background.tex
re

Our previous work on parallel deduplication~\cite{zhangusenix13} is a low-cost
solution, but it is not an inline solution and th backup for all VMs needs to be conducted
around the same time. It does not address  snapshot deletion and file system block 
packaging for fault tolerance.

>>>>

\comments{
Our previous work on parallel deduplication~\cite{zhangusenix13} is a low-cost
solution, but it is not an inline solution and th backup for all VMs needs to be conducted
around the same time. 
It does not address  snapshot deletion and file system block packaging for fault tolerance. 
}

Our previous work on parallel deduplication~\cite{zhangusenix13,WeiZhangIEEE} 
is not a comprehensive VM-centric solution and does not address  snapshot deletion and file system block 
packaging for fault tolerance.  The work in ~\cite{zhangusenix13} a low-cost
solution, but requires all backups for each day to occur at the same time and is less flexible than the approach presented here.

% a low cost solution, but it not an inline solution and 
%the backup for all VMs needs to be conducted
%around the same time.

### 2014-07-19 snapshot_system/paper_conc.tex
re

\section{Conclusion}
\label{sect:conclusion}
%In this paper we propose a collocated backup service built on
%the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The main contribution of this paper is a low-profile and VM-centric deduplication scheme that
uses  a small amount of system resource in both snapshot deduplication and deletion
while delivering competitive deduplication efficiency.
The VM-centric design also allows an optimization for  better fault isolation and tolerance.
Our evaluation has the following results.
\begin{itemize}
\item {\em Competitive deduplication efficiency with a tradeoff}. 
The proposed VC scheme 
can accomplish over  96\% of what complete global
deduplication can do. In comparison, SRB~\cite{Dong2011,extreme_binning09}
can accomplish 97.86\% while 
the sampled index~\cite{Guo2011} can deliver 97\% for a different test dataset. 
\item {\em Lower resource usage in deduplication}. 
The VC scheme achieves a 85K:1 ratio between raw snapshot data and memory usage
and is much more memory-efficient than 
SRB  with 4K:1 ratio and sampled index with 20K:1.
VC with 100 machines takes 1.31 hours to complete the backup of 2,500 VMs
using 19\% of one core and 16.7\% of IO bandwidth per machine. 
Processing time of VC is 23.9\% faster  than SRB in our tested cases
and the per-machine throughput of VC is reasonable based on the result 
from~\cite{Guo2011}. 
Noted that both VC and SRB are designed for a distributed cluster architecture while
the sampled index method is for a single-machine architecture.
\item {\em Lower cost in snapshot deletion}. 
The deletion in VC with summary vectors
takes less than a minute for snapshot deletion using 15MB memory per machine
and all machines can run this operation in parallel without data dependence. 
The grouped mark-sweep method can take 43 hours using
1.2 to 3GB memory per machine in our tested cases.
The leak repair in VC takes 0.82 hours using 50MB on average while processing
some VMs may use upto 1.96GB. Such a repair is conducted infrequently (e.g. every 19.6 days) 
\item {\em Higher availability}. 
%The availability of snapshots increases substantially when
%adding more replication for popular cross-VM file blocks and packaging chunks from the same 
%VM in one file system block. 
The snapshot availability of VC is 99.66\% with 3 failed machines in a 100-machine cluster
while it is 69.54\% for  a VM oblivious approach.
The analysis shows that the replication degree
for the popular data set between 6 and 9 is good enough when the replication degree
for other data blocks is 3, and adds only a small cost to storage.
\end{itemize}
The offline PDS recomputation does require some modest I/O and memory resource and 
since the recomputing frequency is relatively low (e.g. on a monthly basis), we expect such resource
consumption is acceptable in practice.

>>>>

\section{Conclusions}
\label{sect:conclusion}
%In this paper we propose a collocated backup service built on
%the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The main contribution of this paper is the development and comparative study of a low-profile and VM-centric deduplication scheme that
uses  a small amount of system resource in both snapshot deduplication and deletion
while delivering competitive deduplication efficiency.
The VM-centric design also allows an optimization for  better fault isolation and tolerance.
Our evaluation has the following results.
\begin{itemize}
\item {\em Competitive deduplication efficiency with a tradeoff}. 
The proposed VC scheme 
can accomplish over  96\% of what complete global
deduplication can do. In comparison, SRB~\cite{Dong2011,extreme_binning09}
can accomplish 97.86\% while 
the sampled index~\cite{Guo2011} can deliver 97\% for a different test dataset. 
\item {\em Lower resource usage in deduplication}. 
The VC scheme achieves a 85K:1 ratio between raw snapshot data and memory usage
and is much more memory-efficient than 
SRB  with 4K:1 ratio and sampled index with 20K:1.
VC with 100 machines takes 1.31 hours to complete the backup of 2,500 VMs
using 19\% of one core and 16.7\% of IO bandwidth per machine. 
Processing time of VC is 23.9\% faster  than SRB in our tested cases
and the per-machine throughput of VC is reasonable based on the result 
from~\cite{Guo2011}. 
Noted that both VC and SRB are designed for a distributed cluster architecture while
the sampled index method is for a single-machine architecture.
\item {\em Lower cost in snapshot deletion}. 
The deletion in VC with summary vectors
takes less than a minute for snapshot deletion using 15MB memory per machine
and all machines can run this operation in parallel without data dependence. 
The grouped mark-sweep method can take 43 hours using
1.2 to 3GB memory per machine in our tested cases.
The leak repair in VC takes 0.82 hours using 50MB on average while processing
some VMs may use upto 1.96GB. Such a repair is conducted infrequently (e.g. every 19.6 days) 
\item {\em Higher availability}. 
%The availability of snapshots increases substantially when
%adding more replication for popular cross-VM file blocks and packaging chunks from the same 
%VM in one file system block. 
The snapshot availability of VC is 99.66\% with 3 failed machines in a 100-machine cluster
while it is 69.54\% for  SRB or a VM-oblivious approach.
The analysis shows that the replication degree
for the popular data set between 6 and 9 is good enough when the replication degree
for other data blocks is 3, and adds only a small cost to storage.
\end{itemize}
The offline PDS recomputation does require some modest I/O and memory resource and 
since the recomputing frequency is relatively low (e.g. on a monthly basis), we expect such resource
consumption is acceptable in practice.

### 2014-07-19 snapshot_system/paper_dedup.tex
re

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
What  motivates us to use PDS
is that for those chunks that appear in different VMs, the top $k$ most popular items
dominate the appearance distribution.  
Figure~\ref{fig:Datazipf} shows the distribution of chunk popularity for a data trace 
from Alibaba with 2500 VMs discussed in Section ~\ref{sect:evaluation}.
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
%the distribution of chunk popularity~\cite{WeiZhangIEEE} 
%based on a production data trace from Alibaba with over 2500 VMs.
The distribution from this figure is Zipf-like. 
Let  $\sigma$ denote as the percentage of unique data chunks belonging to PDS and 
from the evaluation in Section~\ref{sect:evaluation}, we find that
$\sigma$ with about  2\% can deliver a fairly competitive deduplication efficiency.
%$\sigma$ with about  range of 2 to 4\% can deliver a fairly competitive deduplication efficiency.
%For example,
%with top 2\% of the most popular chunks stored in PDS, up-to 40\% of the raw backup 
%data can be eliminated by fingerprint lookup in PDS index.

>>>>

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
What  motivates us to use PDS
is that for those chunks that appear in different VMs, the top $k$ most popular items
dominate the appearance distribution~\cite{WeiZhangIEEE}.  
Figure~\ref{fig:Datazipf} shows the distribution of chunk popularity for a data trace 
from Alibaba with 2500 VMs discussed in Section ~\ref{sect:evaluation}.
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
%the distribution of chunk popularity~\cite{WeiZhangIEEE} 
%based on a production data trace from Alibaba with over 2500 VMs.
The distribution from this figure is Zipf-like. 
Let  $\sigma$ denote as the percentage of unique data chunks belonging to PDS and 
from the evaluation in Section~\ref{sect:evaluation}, we find that
$\sigma$ with about  2\% can deliver a fairly competitive deduplication efficiency.
%$\sigma$ with about  range of 2 to 4\% can deliver a fairly competitive deduplication efficiency.
%For example,
%with top 2\% of the most popular chunks stored in PDS, up-to 40\% of the raw backup 
%data can be eliminated by fingerprint lookup in PDS index.

### 2014-07-19 snapshot_system/paper_intro.tex
re

%Without deduplication,
Source-side deduplication (e.g. ~\cite{Symantec2009})
 eliminates duplicates before backup data is transmitted
to a secondary storage, which saves network bandwidth significantly; however its resource
usage can impact other co-located cloud services. 
It is memory-intensive to compare a large number of fingerprints and identify
duplicates, even optimization or approximation techniques are developed~\cite{Guo2011,Dong2011,extreme_binning09}. 
When deleting unused snapshots, a grouped mark-and-sweep approach~\cite{Guo2011}  
is effective while it still carries a significant cost, especially in a distributed
setting.  Another side effect of deduplication is the possible loss of failure
resilience~\cite{Reliability06}.  
%Separate files share the same physical copies of blocks that are
%logically duplicated among them.  
If a shared block is lost, all files that
share that block are affected.  
A cloud may offload backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out undeduplicated backup data wastes a huge amount
of network bandwidth that would otherwise be available to user VMs. 
\comments{
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).
}

This paper considers a backup service for converged cloud architectures
that is co-located with user VMs, sharing the cloud compute, network, and
storage resource with them.  We propose a VM-centric approach that
simplifies deduplication and snapshot deletion.
The key advantage of this approach  is that while maintaining  a competitive deduplication efficiency,
the scheme requires little resource usage to minimize the impact on other collocated cloud services.
\comments{
The key contribution of this work is a set of methodologies to address the following two challenges:
1) simple deduplication and snapshot deletion with competitive efficiency and 
a small resource usage without affecting other collocated cloud services;
2) improved fault isolation for the sharing of deduplicated data,
given that node failures in converged architecture are the norm rather than
exception.

>>>>

%Without deduplication,
Source-side deduplication (e.g. ~\cite{Symantec2009})
 eliminates duplicates before backup data is transmitted
to a secondary storage, which saves network bandwidth significantly; however its resource
usage can impact other co-located cloud services. 
It is memory-intensive to compare a large number of fingerprints and identify
duplicates, even optimization or approximation techniques are developed~\cite{Guo2011,Dong2011,extreme_binning09}. 
When deleting unused snapshots, a grouped mark-and-sweep approach~\cite{Guo2011}  
is effective while it still carries a significant cost, especially in a distributed
setting.  Another side effect of deduplication is the possible loss of failure
resilience~\cite{Reliability06}.  
%Separate files share the same physical copies of blocks that are
%logically duplicated among them.  
If a shared block is lost, all files that
share that block are affected.  
A cloud may offload backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out raw and undeduplicated backup data wastes a huge amount
of network bandwidth that would otherwise be available to user VMs. 
\comments{
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).
}

This paper considers a backup service for converged cloud architectures
that is co-located with user VMs, sharing the cloud compute, network, and
storage resource with them.  The main contribution of this paper is
the development and comparative study of  a VM-centric approach that
simplifies deduplication and snapshot deletion and improves snapshot availability.
The key advantage of this approach  is that while maintaining  a competitive deduplication efficiency,
the scheme requires little resource usage to minimize the impact on other collocated cloud services.
\comments{
The key contribution of this work is a set of methodologies to address the following two challenges:
1) simple deduplication and snapshot deletion with competitive efficiency and 
a small resource usage without affecting other collocated cloud services;
2) improved fault isolation for the sharing of deduplicated data,
given that node failures in converged architecture are the norm rather than
exception.
